HUME: Human UCCA-Based Evaluation of Machine Translation

### 2016-03-16 paper/main.tex
ondrej's comments and fixes

\documentclass[11pt]{article}
\usepackage{acl2016}
\usepackage{amsmath}
\usepackage{paralist}
\usepackage{subfig}
\usepackage{times}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{url}
\usepackage{pgfplotstable}
\usepackage{hyperref}
\usepackage{color}
\usepackage{lipsum,adjustbox}

@@

\paragraph{Semantic Representation.}
UCCA (Universal Conceptual Cognitive Annotation, \inparcite{abend2013universal}) is a cross-linguistically applicable, lightweight
scheme for semantic annotation. Formally, a UCCA structure is a directed acyclic graph (DAG),
whose leaves correspond to the words of the text.
The graph’s nodes, called {\sc units}, are either terminals or several elements jointly
viewed as a single entity according to some semantic or cognitive consideration. Edges bear
a category, indicating the role of the sub-unit in the structure the unit
represents. 
UCCA currently focuses on annotating argument structures (adjectival, nominal, verbal and others) and
relations between them. The most basic notion is the {\sc scene}, which describes a movement, an
action or a state which is persistent in time. Each Scene contains one main relation, as well
as one or more participants. For example, the sentence ``After graduating, Tom moved to NYC''
contains two Scenes, whose main relations are ``graduating'' and ``John''\bh{Don't you mean ``graduating'' and
``moved''? And where did ``John'' come from?}. The participant ``John''
is a part of both Scenes, while ``NYC'' only of the latter. Further categories account for
relations between Scenes and the internal structures of complex participants and relations.
See \figref{fig:ucca_example}.

@@

@@

This section summarises the manual annotation procedure performed in order
to compute the HUME measure. 
We mark the source sentence as $s$ and the translation sentence as $t$. 
The procedure involves two manual steps: (1) UCCA-annotating $s$, 
(2) human judgements as to the translation quality of each semantic unit of $s$,
defined by its UCCA annotation, relative to $t$.
We note that UCCA annotation is performed once for every source sentence, irrespective
of the number of its translations we wish to evaluate\footnote{Instead of
relying on the source,
the evaluation could be based on the UCCA structure for the \emph{reference}
translation. This would allow monolingual annotators to take part in the
evaluation. On the downside, the translation would have to be aligned to the
reference and more importantly, we would risk inaccurate judgements due to
the naturally occurring differences between the source and its reference
translations.}.

\paragraph{UCCA Annotation.}
We begin by creating UCCA annotations for the source sentence, following the
UCCA guidelines\footnote{All UCCA-related resources can be found
  here: \url{http://www.cs.huji.ac.il/~oabend/ucca.html}}.
Let $L$ be the set of possible UCCA labels.
Given a sentence $s = (w_1, \ldots,w_n)$, its UCCA annotation is a DAG $G=(V,E,\ell)$,
where $\ell: V \rightarrow L$, and where $W = \{w_1,\ldots,w_n\}$ correspond to the leaves of $G$.
For every node $v \in V$, we define its {\it yield} $yld(v)$ as $v$'s descendants in $W$.
The semantic units for a sentence $s$ induced by the annotation $G$,
is the set of yields of nodes in $V$. %We denote it with $Units(s)$ for brevity.
We note that the translation is not UCCA-annotated.

\begin{figure*}[t]
    \includegraphics[width=1\textwidth]{ucca-tree-mteval}
    \caption{UCCA Tree evaluated in comparison to an aligned ``translation'' (an
	English paraphrase).}
    \label{fig:ucca-tree-hume}
\end{figure*}

@@

Human annotation of the semantic units requires first deciding whether
a unit is {\it atomic}, i.e., that it has no meaning-bearing
sub-units, or {\it structural}. In most cases, atomic units 
correspond to individual words, but they may also correspond to unanalyzable
multi-word expressions, i.e., expressions that do not have meaningful sub-units.
When a multi-word unit is labeled as atomic, all its sub-units' annotations are discarded
from the evaluation.

@@

Structural units have sub-units (children in the UCCA graph), which may either be
atomic or structural, and represent the relation between the sub-units.
For instance, a Scene unit represents the relation between the
Scene elements, such as the predicate (main relation), participants and adverbials.
Structural units are labeled as ``Adequate'' or ``Bad'', meaning
that the relation between the sub-units went wrong.
We make a three-way distinction between translation qualities with structural
nodes, as opposed to the two-way distinction with atomic units, as lexical units
often can have slight errors, in which cases we do want to provide partial credit.

@@

\XXX{Could someone review this for Ondrej.}
Note that for non-configurational languages, the relationship between the
governing and the dependent units is often expressed using morphological
properties of the dependents instead of their ordering (this holds especially for the verb and its
modifiers, less so for components of noun phrases). Since errors in morphology
should be expressed on the atomic units, we would expect
a higher proportion of ``Adequate'' structural units for Czech and Polish
compared to German and Romanian\oa{is this because that error would already be marked
  in the atomic units, and thus not in the structural ones?}.

We defined the labels to reflect adequacy,
rather than fluency judgements. Specifically, annotators are instructed to
label a unit as Adequate if its translation is understandable and preserves
the meaning of the source unit, even if its fluency is impaired.

@@

Another example is given in \figref{fig:ucca-tree-hume}.
%``Golf became an obsession for his oldest daughter, she took daily lessons and became very good'',
The figure shows an annotation of a UCCA tree, and an example annotation for
each of the units.
We see that ``became a passion'' is deemed atomic, as it is translated as a single unit
in $t$, and thus its sub-units, ``became'', ``a'' and ``passion'' are not evaluated
individually.\oa{I'm not sure I agree with the UCCA annotation;
also, translation annotation for ``she'' is missing. and for ``very good'' is wrong}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Annotation Interface}

We turn to discuss the interface for collecting human judgements on the semantic units of the source,
relating to the translation. UCCA annotations of the source are done offline, using the UCCA web
application\footnote{See UCCA webpage.}.

\figref{fig:interface} presents the user interface of HUME. The sentence at the top
shows the complete MT system output. The source sentence appears at the top, and
the source translation directly below it. Underneath the source sentence appears
its list of semantic units. For each semantic unit, including the one for the
entire source sentence, there is a textbox, displaying the yield of that unit.
Information on the category of the unit (e.g., whether it is a Scene, participant etc.).
The user is asked to select a label for each unit,
by clicking the ``A'', ``B'', Green, Orange, or Red buttons to the left of the unit's box.

The interface presents, for each unit, the segment of the translation which aligns to it.
This allows the user, especially in long sentences, to focus her attention on the parts
mostly likely to be relevant for her judgement. As the alignments are automatically derived,
and therefore noisy, the annotator is instructed to treat the aligned text is a cue, but to ignore
it and make a judgement according to the full translation sentence, in cases where the alignment
is misleading.

Concretely, let $s = (w_1,\ldots,w_n)$ be a source sentence, $t = (w'_1,\ldots,w'_n)$ a translation,
and $A \subset 2^{s \times t}$ a many-to-many word alignment between them.
If $u$ is a semantic unit in $s$, whose yield is $yld(u)$, we define the aligned text in
$t$ to be $a(u)=\bigcup_{(x_s,x_t) \in A \wedge x_s \cap yld(u) \neq \emptyset} x_t \subset t$.
In the event that $a(u)$ is discontiguous in the text, we mark words between the left
and right boundaries of $a(u)$ which are not contained in it (intervening words)
in a smaller red font. Intervening words are likely change the meaning of the translation
of $u$, and thus should be paid attention to when considering whether the translation
is correct or not.

For example, consider the sentence $s=$``The man watched the dog'' and the German
translation ``t=Der Mann sieht den Hund an''. Then ``watched'' will be aligned to
``sieht ... an'', and ``den Hund'' will be marked as intervening word. Another example
can be seen in \figref{fig:interface}, where ``ongoing pregnancy'' is translated to
``Schwangerschaft ... laufenden'' (lit. ``pregnancy ... ongoing'').

% Ondrej did the following in a footnote in Section 3.1 Annotation Procedure
%\bh{Somewhere we could explain that we
%could do a similar annotation with monolinguals, although it entails aligning
%reference and MT output}

%There are cases where we cannot usefully compare individual words in the source to words in the target.
%We often need to compare translation at the phrase level. If this is the case, then we can 
% treat the structural node as a lexical unit, and mark it with the traffic light labels. This 
%means that none of the children of this node will be examined in order determine the semantic score
%for this sentence. 
%In Figure~\ref{ucca-tree-mteval}, we can see that the phrase ``became a passion'' has been labelled as a lexical
%node and it has been evaluated as ``Orange'', or partially correct, as the translation of this phrase ``loved'' 
%only partially capture the semantics of the source.

\begin{figure*}[t]
    \begin{center}
    \includegraphics[width=1\textwidth]{hume_interface.png}
    \caption{The HUME annotation tool, based on the UCCA tool. The faded orange box at the top
      presents the translation, the large box below it presents the source. The tree of semantic units
      for the source is given underneath, where the aligned text appears in italics, and intervening words
      appear in red with a smaller font (e.g., in the first box).}
    \label{fig:interface}
    \end{center}
\end{figure*}

\com{
In Figure~\ref{mttool} you we show the MT evaluation tool. The sentence at the top shows the complete MT system output. Underneath the MT output is the  source sentence.  Underneath the source sentence we see its expandable semantic tree structure with both lexical and structural nodes. Lexical nodes only have traffic light annotation, whereas a structural node would normally be labelled ``A'' or ``B'' but could also be labelled as a lexical node, in the case where there is no word to word correspondence for the translation of its children.

 The child components of a structural node are marked with different coloured rectangles. When navigating through the different source sentence nodes, one can see the relevant sections of the translation because the aligned words in the translation are highlighted in the complete sentence above. Aligned translations are also shown in black alongside the source node. If the node is aligned to a set of discontiguous words in the translation, then the unaligned words that appear in between the aligned words are shown in red. Even if these words are not directly aligned to the source node, they 
will likely change the meaning of the translation and must be considered when marking 
the translation as correct or not. The alignments are meant just as a guide. The annotator should  look at the complete translation when deciding on the evaluation of a node. The translation could have content added before or after the node which changes its
 structure or meaning. If an extra component is prepended to a structural node, for example, it should be marked as ``Bad''.
}

@@

@@

@@

\subsection{Copied from emails}

+ allows for multi-word and even non-contiguous elements
+ allows to indicate the quality of the predicates ("Actions")
+/- expects to cover the whole hypothesis with annotations, whereas HMEANT 'gives up' when the
hypothesis is not intelligible, giving a zero score to the bad frame/whole sentence.
This can be a good decision for HMEANT, since our way may artificially boost the chances of
understanding the sentence.

HMEAN disadvantages:
. Importantly: it disregards copula clauses, nominalizations (which are often translated to verbal clauses), inter-clause relations (discourse markers, complement clauses etc.). UCCA-Eval addresses these.

>>>>

\documentclass[11pt]{article}
\usepackage{acl2016}
\usepackage{amsmath}
\usepackage{paralist}
\usepackage{subfig}
\usepackage{times}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{url}
%\usepackage{subfigure}
\usepackage{pgfplotstable}
\usepackage{hyperref}
\usepackage{color}
\usepackage{lipsum,adjustbox}

@@

\paragraph{Semantic Representation.}
UCCA (Universal Conceptual Cognitive Annotation, \inparcite{abend2013universal}) is a cross-linguistically applicable, lightweight
scheme for semantic annotation. Formally, a UCCA structure is a directed acyclic graph (DAG),
whose leaves correspond to the words of the text.
The graph’s nodes, called {\sc units}, are either terminals or several elements jointly
viewed as a single entity according to some semantic or cognitive consideration. Edges bear
a category, indicating the role of the sub-unit in the relation the unit
represents. \XXX{Ondrej suggests to avoid the word 'relation' here. I usually use the
term 'relation' for what is called 'role' here, so I got confused for a while.}
UCCA
currently focuses on annotating argument structures (adjectival, nominal, verbal and others) and
relations between them. The most basic notion is the {\sc scene}, which describes a movement, an
action or a state which is persistent in time. Each Scene contains one main relation, as well
as one or more participants. For example, the sentence ``After graduating, Tom moved to NYC''
contains two Scenes, whose main relations are ``graduating'' and ``John''\bh{Don't you mean ``graduating'' and
``moved''? And where did ``John'' come from?}. The participant ``John''
is a part of both Scenes, while ``NYC'' only of the latter. Further categories account for
relations between Scenes and the internal structures of complex participants and relations.
See \figref{fig:ucca_example}.

@@

%We do not distinguish fluency and adequacy judgements, as they are correlated anyway \cite{ccb}.

@@

This section summarises the procedure for annotation involved in computing HUME.
We mark the source sentence as $s$ and the translation sentence as $t$. 
The procedure involves two manual steps: (1) UCCA-annotating $s$, 
(2) human judgements for each semantic unit of $s$,
defined by its UCCA annotation, of whether its meaning has been preserved in $t$.
We note that (1) is performed once for every source sentence, irrespective
of the number of its translations we wish to evaluate.\footnote{Instead of
relying on the source,
the evaluation could be based on the UCCA structure for the \emph{reference}
translation. This would allow monolingual annotators to take part in the
evaluation. On the downside, the translation would have to be aligned to the
reference and more importantly, we would risk inaccurate judgements due to
the naturally occurring differences between the source and its reference
translations.}

\paragraph{UCCA Annotation.}
We begin by creating UCCA annotations for the source sentence, following the
UCCA guidelines\footnote{All UCCA-related resources can be found
  here: \url{http://www.cs.huji.ac.il/~oabend/ucca.html}}.
Let $L$ be the set of possible UCCA labels.
Given a sentence $s = (w_1, \ldots,w_n)$, the UCCA annotation is a DAG $G=(V,E,\ell)$,
where $\ell: V \rightarrow L$, and where $W = \{w_1,\ldots,w_n\}$ correspond to the leaves of $G$.
For every node $v \in V$, we define its {\it yield} $yld(v)$ as $v$'s descendants in $W$.
The semantic units for a sentence $s$, induced by the annotation $G$
is the set of yields of nodes in $V$. We denote it with $\Units(s)$ for brevity.
We note that the translation need not be annotated.

@@

Human annotation of the semantic units requires first deciding whether
a unit is an {\it atomic unit}, i.e., that it has no meaning-bearing
sub-units, or a {\it structural unit}. In most cases, atomic units would
correspond to individual words, but they may also correspond to opaque
multi-word expressions, i.e., expressions that do not have meaningful sub-units.
When a multi-word unit is labeled as atomic, all its sub-units' annotations are discarded
from the evaluation.

@@

Structural units have sub-units (children in the UCCA graph), which may either be
atomic or structural, and represent the relation between the sub-units.
For instance, a Scene unit 
represents the relation between the main relation, participants and secondary
relations in it\XXX{again, could we avoid the word ``relation'' here? Would
``event'' be an acceptable synonym?}. Structural units are labeled as ``Adequate'' or ``Bad'', meaning
that the relation between the sub-units went wrong.

@@

\XXX{Could someone review this for Ondrej:}
Note that for non-configurational languages, the relationship between the
governing and the dependent units is often expressed using morphological
properties of the dependents instead of their ordering (this holds especially for the verb and its
modifiers, less so for components of noun phrases). Since errors in morphology
should be expressed using the traffic lights on the dependent nodes, we would expect
a higher proportion of ``Adequate'' structural units for Czech and Polish
compared to German and Romanian.

We defined the labels to reflect more adequacy,
rather than fluency judgements. Specifically, annotators are instructed to
label a unit as Adequate if its translation is understandable and preserves
the meaning of the source unit, even if its fluency is impaired.

@@

@@

%composite score (uniform, sub-sets, weighted rather not -- according to categories [or empirically according to agreement])

\subsection{Annotation Interface}

As translations may be long, we provide assistance to the annotator.
Where the annotator is asked to judge the translation quality of a unit $u$,
the part of by highlighting the part of the translation which is likely to be
most relevant for the semantic unit in question.
This is done by highlighting the words in the translation which are aligned

using automatically derived word-to-word  in order to project
this annotation to the MT output.

looking at the translation via the word alignments, and marking
on the source structure which parts have been correctly translated.
There are two kinds of components (nodes): a word or basic semantic unit, and a structural component
which contains one or more sub-components.

In order 
, and use native speakers of the target
language to annotate this output.

Since these annotators have to judge the quality of the translation,
they also needed to be fluent in English.
% Ondrej did the following in a footnote in Section 3.1 Annotation Procedure
%\bh{Somewhere we could explain that we
%could do a similar annotation with monolinguals, although it entails aligning
%reference and MT output}

If any of these changes have occurred and this damages the meaning of the aligned translation, then we mark the 
structural node as ``Bad''. 
However if the arrangement of the child nodes is essentially correct, except that one or
more of the child nodes has themselves been translated wrongly (and so is marked
orange or red), then the structural node should be 
annotated as ``Acceptable''.

\begin{figure*}[t]
    \includegraphics[width=1\textwidth]{ucca-tree-mteval}
    \caption{UCCA Tree evaluated in comparison to an aligned ``translation'' (an
	English paraphrase).}
    \label{ucca-tree-mteval}
\end{figure*}

%TODO describe figure 2 here

There are cases where we cannot usefully compare individual words in the source to words in the target.
We often need to compare translation at the phrase level. If this is the case, then we can 
 treat the structural node as a lexical unit, and mark it with the traffic light labels. This 
means that none of the children of this node will be examined in order determine the semantic score
for this sentence. 
In Figure~\ref{ucca-tree-mteval}, we can see that the phrase ``became a passion'' has been labelled as a lexical
node and it has been evaluated as ``Orange'', or partially correct, as the translation of this phrase ``loved'' 
only partially capture the semantics of the source.

We are separating lexical and structural evaluation in order to simplify evaluation and to localise errors
to their point of origin. For this reason, if a word has been translated incorrectly,
the parent node should still be correct if the number and relationship between its children are the same
as in the source. 
In Figure~\ref{ucca-tree-mteval}, we can see that even though an important child of the first scene is 
incorrectly translated (``Golf''), and the children of are ordered differently, 
the scene itself is marked as ``Acceptable''. The translation of the scene retains the structural 
relationships between the children.
\\

\begin{figure*}[t]
    \begin{center}
    \includegraphics[width=1\textwidth]{interface-annotations}
    \caption{The HUME annotation tool, based on the UCCA tool}
    \label{mttool}
    \end{center}
\end{figure*}

In Figure~\ref{mttool} you we show the MT evaluation tool. The sentence at the top shows the complete MT system output. Underneath the MT output is the  source sentence.  Underneath the source sentence we see its expandable semantic tree structure with both lexical and structural nodes. Lexical nodes only have traffic light annotation, whereas a structural node would normally be labelled ``A'' or ``B'' but could also be labelled as a lexical node, in the case where there is no word to word correspondence for the translation of its children.

 The child components of a structural node are marked with different coloured rectangles. When navigating through the different source sentence nodes, one can see the relevant sections of the translation because the aligned words in the translation are highlighted in the complete sentence above. Aligned translations are also shown in black alongside the source node. If the node is aligned to a set of discontiguous words in the translation, then the unaligned words that appear in between the aligned words are shown in red. Even if these words are not directly aligned to the source node, they 
will likely change the meaning of the translation and must be considered when marking 
the translation as correct or not. The alignments are meant just as a guide. The annotator should  look at the complete translation when deciding on the evaluation of a node. The translation could have content added before or after the node which changes its
 structure or meaning. If an extra component is prepended to a structural node, for example, it should be marked as ``Bad''.

@@

\subsection{One vs. Two Structures}

\subsection{Verbal Structures Only?}

\subsection{Monolingual vs. Bilingual Evaluation}
+/- expects to cover the whole hypothesis with annotations, whereas HMEANT 'gives up' when the
hypothesis is not intelligible, giving a zero score to the bad frame/whole sentence.
This can be a good decision for HMEANT, since our way may artificially boost the chances of
understanding the sentence.

\subsection{Technical Differences}

+ allows for multi-word and even non-contiguous elements

+ allows to indicate the quality of the predicates ("Actions")

\subsection{XXX to be moved}

@@

\subsection{from e-mails}

@@
